---
title: "Where Surfers travel"
author: "by [Yan Holtz](https://github.com/holtzy/) - `r format(Sys.time(), '%d %B %Y')`"
output:
  epuRate::epurate:
    toc: FALSE
    code_folding: "show"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
```{r echo=FALSE}
# Just to add space between the introduction of the document
knitr::asis_output(htmltools::htmlPreserve("<br><br>"))
```


> I find it not too bad to start with a couple of sentences of introduction. Explain briefly what this document is going to talk about. May be add a [useful link](www.r-graph-gallery.com) relative to this project

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(jpeg)
library(maps)
library(geosphere)
library(grid)

```



# Get the data
***

##NASA image
Using the NASA images, inspired from [here](https://www.dataplanes.org/notes/2018/01/27/flight-routes-night-lights). 
```{r}
# Download NASA night lights image
download.file("https://www.nasa.gov/specials/blackmarble/2016/globalmaps/BlackMarble_2016_01deg.jpg", destfile = "DATA/BlackMarble_2016_01deg.jpg", mode = "wb")

# Load picture and render
earth <- readJPEG("DATA/BlackMarble_2016_01deg.jpg", native = TRUE)
earth <- rasterGrob(earth, interpolate = TRUE)
```

##Surfer Data
```{r}
load("DATA/Surf_Hashtag_Data.Rdata")

# Keep tweets with origin localization, and current localization
travel_data=data[!is.na(data$travellon) & !is.na(data$homelon) , ]
dim(travel_data)

# Count how many times we have each unique connexion + order by importance
summary=travel_data %>% 
  count(homelat,homelon,homecontinent, travellat,travellon,travelcontinent) %>%
  #filter(n>1) %>%
  arrange(n) 
dim(summary)

# A function that makes a dateframe per connection (we will use these connections to plot each lines)
data_for_connection=function( dep_lon, dep_lat, arr_lon, arr_lat, group){
  inter <- gcIntermediate(c(dep_lon, dep_lat), c(arr_lon, arr_lat), n=50, addStartEnd=TRUE, breakAtDateLine=F)             
  inter=data.frame(inter)
  inter$group=NA
  diff_of_lon=abs(dep_lon) + abs(arr_lon)
  if(diff_of_lon > 180){
    inter$group[ which(inter$lon>=0)]=paste(group, "A",sep="")
    inter$group[ which(inter$lon<0)]=paste(group, "B",sep="")
  }else{
    inter$group=group
  }
  return(inter)
}

# Création d'un dataframe complet avec les points de toutes les lignes à faire.
data_ready_plot=data.frame()
for(i in c(1:nrow(summary))){
  tmp=data_for_connection(summary$homelon[i], summary$homelat[i], summary$travellon[i], summary$travellat[i] , i)
  tmp$homecontinent=summary$homecontinent[i]
  tmp$n=summary$n[i]
  data_ready_plot=rbind(data_ready_plot, tmp)
}
dim(data_ready_plot)
data_ready_plot$homecontinent=factor(data_ready_plot$homecontinent, levels=c("Asia","Europe","Australia","Africa","North America","South America","Antarctica"))

```



#- Make the plot

```{r}
ggplot() + 
  annotation_custom(earth, xmax = 180, ymin = -Inf, ymax = Inf) +
  geom_line(data=data_ready_plot, size=0.5, aes(x=lon, y=lat, group=group, colour=homecontinent, alpha=n)) +
  xlim(-170, 185)+ ylim(-59, 80) +
  theme(
    legend.position="none",
    #plot.margin = unit(rep(-1.25,4),"lines"),
    plot.background = element_rect(fill = "#00001C"),
    panel.background = element_rect(fill='#00001C',colour='#00001C' ),
    panel.grid=element_blank(), 
    panel.border = element_blank(),
    axis.title=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank()
  ) 
ggsave("IMG/Where_Surfers_Travel.png", width = 36, height = 18, units = "in", dpi = 100)
```






ggplot() +
  annotation_custom(earth) +
  geom_point(aes(x=0, y=0), size=8, color="yellow") 
ggsave("IMG/Where_Surfers_Travel.png", width = 36, height = 18, units = "in", dpi = 100)
```


















```{r}
# Download OpenFlights data
download.file("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airlines.dat",
destfile = "airlines.dat", mode = "wb")
download.file("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat", 
destfile = "airports.dat", mode = "wb")
download.file("https://raw.githubusercontent.com/jpatokal/openflights/master/data/routes.dat", 
destfile = "routes.dat", mode = "wb")

# Import data
library(data.table)
airlines <- fread("airlines.dat", sep = ",", skip = 1)
airports <- fread("airports.dat", sep = ",")
routes <- fread("routes.dat", sep = ",")
head(airlines)



# Add column names
colnames(airlines) <- c("airline_id", "name", "alias", "iata", "icao", "callsign", "country", "active")
colnames(airports) <- c("airport_id", "name", "city", "country", "iata", "icao", "latitude", "longitude", 
"altitude", "timezone", "dst", "tz_database_time_zone", "type", "source")
colnames(routes) <- c("airline", "airline_id", "source_airport", "source_airport_id", "destination_airport", 
"destination_airport_id", "codeshare", "stops", "equipment")

# Convert character to numeric
routes$airline_id <- as.numeric(routes$airline_id)

# Merge airline data with data on routes
flights <- left_join(routes, airlines, by = "airline_id")

# Merge data on flights with information on airports
airports_orig <- airports[, c(5, 7, 8)]
colnames(airports_orig) <- c("source_airport", "source_airport_lat", "source_airport_long")

airports_dest <- airports[, c(5, 7, 8)]
colnames(airports_dest) <- c("destination_airport", "destination_airport_lat", "destination_airport_long")

flights <- left_join(flights, airports_orig, by = "source_airport")
flights <- left_join(flights, airports_dest, by = "destination_airport")

# Remove observations with missing values
flights <- na.omit(flights, cols = c("source_airport_long", "source_airport_lat", "destination_airport_long", "destination_airport_lat"))






# GCINTERMEDIATE
# Split the data into separate data sets
flights_split <- split(flights, flights$name)

# Calculate intermediate points between each two locations
flights_all <- lapply(flights_split, function(x) gcIntermediate(x[, c("source_airport_long", "source_airport_lat")], 
x[, c("destination_airport_long", "destination_airport_lat")], 
100, breakAtDateLine = FALSE, addStartEnd = TRUE, sp = TRUE))

# Turn data into a data frame for mapping with ggplot2
library(plyr)
flights_fortified <- lapply(flights_all, function(x) ldply(x@lines, fortify))

# Unsplit lists
flights_fortified <- do.call("rbind", flights_fortified)

# Add and clean column with airline names
flights_fortified$name <- rownames(flights_fortified)
flights_fortified$name <- gsub("\\..*", "", flights_fortified$name)

# Extract first and last observations for plotting source and destination points (i.e., airports)
flights_points <- flights_fortified %>%
  group_by(group) %>%
  filter(row_number() == 1 | row_number() == n())



ggplot() +
  annotation_custom(earth, xmin = -180, xmax = 180, ymin = -90, ymax = 90) +
  geom_path(aes(long, lat, group = id, color = name), alpha = 0.0, size = 0.0, data = flights_fortified) + 
  geom_path(aes(long, lat, group = id, color = name), alpha = 0.2, size = 0.3, color = "#f9ba00", data = flights_fortified[flights_fortified$name == "Lufthansa", ]) + 
  geom_point(data = flights_points[flights_points$name == "Lufthansa", ], aes(long, lat), alpha = 0.8, size = 0.1, colour = "white") +
  theme(panel.background = element_rect(fill = "#05050f", colour = "#05050f"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks.length = unit(0, "cm"),
        legend.position = "none") +
  annotate("text", x = -150, y = -18, hjust = 0, size = 14,
           label = paste("Lufthansa"), color = "#f9ba00", family = "Helvetica Black") +
  annotate("text", x = -150, y = -26, hjust = 0, size = 8, 
           label = paste("Flight routes"), color = "white") +
  annotate("text", x = -150, y = -30, hjust = 0, size = 7, 
           label = paste("lhehnke.github.io || NASA.gov || OpenFlights.org"), color = "white", alpha = 0.5) +
  coord_equal()



```
